\begin{resume}

  \resumeitem{个人简历}

  王腾飞，男，汉族，1990年2月出生于河南省襄城县。

  2007年9月，考入同济大学海洋与地球科学学院。

  2011年6月获地球物理学学士学位。

  2011年9月，免试直接攻读同济大学固体地球物理学博士学位至今。

  \resumeitem{已发表期刊论文} % 发表的和录用的合在一起

  \begin{enumerate}[{[}1{]}]
	  \item {\bf{Tengfei Wang}}  and Jiubing Cheng, 2017, Elastic full waveform inversion
		  based on mode decomposition: the approach and mechanism. 
		  \emph{Geophysical Journal International}, 209, 606–622.
	  \item
		  程玖兵，陈茂根，{\bf{王腾飞}}，康玮，2014，各向异性介质qP波传播描述II：分离纯模式标量波. \emph{地球物理学报},
	  57(10)，3389-3401.
  \item  程玖兵，康玮，{\bf{王腾飞}}, 2013,
  各向异性介质qP波传播描述I：伪纯模式波动方程. \emph{地球物理学报},
	  56(10)， 3474-3486.
  \item  Cheng Jiubing, {\bf{Wang Tengfei}}, Wang Chenlong, and Geng Jianhua, 2012, Azimuth-preserved local
	  angle-domain prestack time migration in isotropic, vertical transversely isotropic and
	  azimuthally anisotropic media. \emph{Geophysics}, 77, 2, S51-S64.
  \end{enumerate}

  \resumeitem{已发表会议论文} % 有就写，没有就删除
  \begin{enumerate}[{[}1{]}]
	  \item {\bf{Wang Tengfei}}, Jiubing Cheng and Chenlong Wang, 2017, Elastic wave-equation reflection
		  traveltime inversion using dynamic warping and wave mode decomposition. \emph{79th EAGE
		  Conference \& Exhibition 2017}, Paris, France.
	  \item {\bf{Wang Tengfei}} and Jiubing Cheng, 2016, Comparison between mode decomposition based and
		  Gaussian Newton methods in elastic full waveform inversion. \emph{ 78th EAGE Conference \&
		  Exhibition}, Vienna, Austria.
	  \item Yang T, {\bf{Wang T}}, Cheng J, 2016, Regularization method of tomography based on the ADCIGs
		  with geologic dip information. \emph{SEG International Geophysical
		  Conference} Beijing, China.
	  \item {\bf{Wang Tengfei}}, Jiubing Cheng and Chenlong Wang, 2015, Elastic wave mode decoupling for
		  full waveform inversion. \emph{77th EAGE Conference \& Exhibition}, Mardrid, Spain.
	  \item {\bf{Wang Tengfei}} and Jiubing Cheng, 2015, Elastic wave mode decoupling for full waveform
		  inversion. \emph{85th SEG Annual International Meeting}, Expanded Abstracts.
	  \item   {\bf{Wang T}}, Kang W, Cheng J, 2014, Migration velocity model building using local angle domain non-
		  linear tomography. \emph{International Geophysical Conference and Exposition}, 
		  Beijing, China. 
	  \item Wang Chenlong, Cheng Jiubing and {\bf{Wang Tengfei}}, 2014, Local angle domain elastic reverse
		  time migration in TI media. \emph{76th EAGE Conference \& Exhibition}, Amsterdam,
		  Netherlands. 
	  \item {\bf{Wang Tengfei}}, Cheng Jiubing and Kang Wei, 2012, Pure mode modeling and reverse-time
		  migration of P-wave in HTI and orthorhombic media.\emph{82nd Ann.
		  Internat Mtg. SEG}, Expanded Abstracts, Las Vegas, USA.
			  
  \end{enumerate}
\end{resume}
